Indicators for Ecosystem Condition in Norway

### 2022-10-07 naturtype2.Rmd
new plot

```{r, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggplot2)
library(data.table)
library(stringr)
library(units)
library(dplyr)
library(tidyr)
knitr::opts_chunk$set(echo = TRUE)
```

@@

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Slåttemark"                                         ] <- "T32 | partial"

@@

#### Add number of location and total area mapped
```{r}
dat2$km2 <- units::drop_units(dat2$area/10^6)
num <- aggregate(data = dat2,
              km2~naturtype,
              FUN = sum)

num$numberOfLocalities <- aggregate(data = dat2,
              km2~naturtype,
              FUN = length)[,2]
#num <- num[num$naturtype %in% ntypDF2$Nature_type,]

@@

@@

```{r, fig.cap="The total number of mapped localities for nature types associated with three selected main ecosystems"}
ntyp_fill %>%
  arrange(numberOfLocalities) %>%
  mutate(Nature_type=factor(Nature_type, levels=Nature_type)) %>%   # This trick update the factor levels
  ggplot( aes(x = Nature_type,
               y = numberOfLocalities)) +
    geom_segment( aes(xend=Nature_type, yend=0)) +
    geom_point( size=4, aes(group = Ecosystem,
                            colour= Ecosystem)) +
    coord_flip() +
    theme_bw() +
    xlab("")+
    ylab("Number of localities")+
  theme(legend.position = c(0.55, 0.25),
        legend.background = element_rect(fill = "white", color = "black"))

@@

>>>>

```{r, include=FALSE}
library(knitr)
library(sf)
library(tmap)
library(ggplot2)
library(data.table)
library(stringr)
library(units)
library(dplyr)
library(RColorBrewer)
library(forcats)
library(tidyr)
knitr::opts_chunk$set(echo = TRUE)
```

@@

ntypDF2$hovedgruppe[ntypDF2$Nature_type == "Slåttemark"                                         ] <- "T32 | special management"

@@

#### Add number of location and total area mapped
```{r}
dat2$km2 <- units::drop_units(dat2$area/10^6)

num <- as.data.frame(dat2) %>%
  group_by(naturtype)%>%
  summarise(km2 = sum(km2),
            numberOfLocalities = n())

@@

## Plots

@@

### Area per NiN main type conditioned on ecosystem and adding empty main types
First I will add the NiN main types that are not covered by any nature types.

```{r}
open <- c(
  "T1",
  "T2",
  "T6",
  "T8",
  "T11",
  "T12",
  "T13",
  "T15",
  "T16",
  "T17",
  "T18",
  "T20",
  "T21",
  "T23",
  "T24",
  "T25",
  "T26",
  "T27",
  "T29"
)

semi <- c(
  "T31",
  "T32",
  "T33",
  "T34",
  "T40",  # are these included
  "T41"   # are these included
)

wetland <- c(
  "V1",
  "V3",
  "V4",
  "V6",
  "V9",
  "V10",
  "L4"    # this will probably not be part of the assessments

)

@@

Adding the non-mapped open types
```{r}
open[open %!in% ntyp_fill$NiN_mainType]
```

```{r}
ntyp_fill2 <- ntyp_fill %>%
  add_row(NiN_mainType = "T6", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0) %>%
  add_row(NiN_mainType = "T11", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0) %>%
  add_row(NiN_mainType = "T13", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0) %>%
  add_row(NiN_mainType = "T16", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0) %>%
  add_row(NiN_mainType = "T23", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0) %>%
  add_row(NiN_mainType = "T24", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0) %>%
  add_row(NiN_mainType = "T25", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0) %>%
  add_row(NiN_mainType = "T26", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0) %>%
  add_row(NiN_mainType = "T27", Ecosystem = "Naturlig åpne områder under skoggrensa", NiN_mainTypeCoverage = "Not mapped", km2 = 0)

Adding the non-mapped semi-natural types
```{r}
semi[semi %!in% ntyp_fill$NiN_mainType]
```
```{r}
ntyp_fill2 <- ntyp_fill %>%
  add_row(NiN_mainType = "T40", Ecosystem = "Semi-naturlig mark", NiN_mainTypeCoverage = "Not mapped", km2 = 0)
```

Adding the non-mapped wetland types
```{r}
wetland[wetland %!in% ntyp_fill$NiN_mainType]
```
This is Våtsnøleie, and my guess is that it IS mapped, but grouped with the aloine ecosystem. I'll therefore not add it in here, but mention it in the figure caption below.

```{r, fig.cap="The areas mapped for each NiN main type, and the degree of spatial representativity (or coverage) of the mapping units (nature types). V6 Våtsnøleie og snøleiekilde is not included."}
mySize <- 10
ntyp_fill2 %>%
  # combining two classes to get a better colour palette
  mutate(NiN_mainTypeCoverage = 
           replace(NiN_mainTypeCoverage, NiN_mainTypeCoverage == "southern", "partial"),
         NiN_mainTypeCoverage = 
           replace(NiN_mainTypeCoverage, NiN_mainTypeCoverage == "calcareous and dry", "calcareous")) %>%

  group_by(NiN_mainType, NiN_mainTypeCoverage) %>%
  mutate(km2 = sum(km2)) %>%
  slice_head()%>%
  select(Ecosystem,
         NiN_mainType,
         NiN_mainTypeCoverage,
         km2)%>%
  ungroup() %>%
  mutate(NiN_mainType = fct_reorder(NiN_mainType, km2)) %>%
  ggplot(aes(x = NiN_mainType,
             y = km2,
             fill = NiN_mainTypeCoverage))+
    geom_bar(stat = "identity",
             colour = "grey40",
             position = "stack")+
    theme_bw(base_size = mySize)+
    coord_flip()+
  labs(x = "NiN main type",
       y = expression(km^2),
       fill = "Spatial coverage")+
  scale_fill_brewer(palette = "Set1")+
  facet_wrap(vars(Ecosystem),
             scales = "free",
             labeller = label_wrap_gen(width=25))

### For each NiN main type, the proportion mapped with each NiN variable

### For each ecosystem, the proportion mapped with each NiN variable
